Analyzing the Privacy Paradox Using a Nationally Representative Three-Wave Panel Study

### 2021-03-05 manuscript.RmD
- Updates for Revision 4

However, current research on the privacy paradox has one major limitation. 
To the best of our knowledge, most empirical studies conducted so far have investigated the privacy paradox from a between-person perspective.
By employing empirical tests of relations between people (e.g., cross-sectional questionnaires analyzed with multiple regression or Pearson correlations), studies have analyzed whether people who are more concerned than _others_ also share less personal information than _others_.
Although such a perspective is interesting and represents a viable first step, it cannot make informed claims regarding causality.
The privacy paradox, however, implies a causal perspective:
Does a person, if _he or she_ becomes more concerned about online privacy, then also share less personal information?
This mismatch is problematic because although between-person relations are, except for some special cases, a _necessary_ condition for causal within-person effects, they are by no means a _sufficient_ one.
For example, it could be that the between-person relation is determined other third variables. 
Hence, as the next step in investigating the privacy paradox and to better understand the causal relation between privacy concerns and information sharing, it is necessary to conduct studies with within-person designs.

@@

Specifically, two of the most influential factors that affect online information sharing are (a) strong subjective norms [@heirmanPredictingAdolescentsDisclosure2013] and (b) expected benefits [@krasnovaOnlineSocialNetworks2010]. 
In other words, users often prioritize social support, special offers, or improved services, accepting that their privacy will be diminished.
Sometimes, privacy concerns do not relate information sharing, because users lack the skills, knowledge, or literacy to change their online behavior, creating feelings of apathy or cynicism [@hargittaiWhatCanReally2016; @hoffmannPrivacyCynicismNew2016].
Likewise, personal information is also often shared by others, a phenomenon described as "networked privacy" [@marwickNetworkedPrivacyHow2014], which further reduces the power of individuals to determine how much personal information can be found online.
@trepteRiskyBehaviorsHow2014 listed several factors that can additionally attenuate the relation:
lack of strength of concerns,
absence of negative personal experiences, 
or situational constraints due to social desirability.
In conclusion, also in the context of the privacy paradox it _not_ reasonable to expect a _perfect_ relation between attitudes and behaviors.
However, we should still expect to find a relation that is _small_ or _moderate_.

@@

It has been argued that privacy attitudes could bridge the gap between concerns and information sharing [e.g., @dienlinPrivacyParadoxRelic2015]. 
In contrast to privacy concerns, privacy attitudes capture a more explicit, fluctuating cognitive appraisal [@tsay-vogelSocialMediaCultivating2018]. 
Although both variables are related to information disclosure, attitudes are likely the better predictor. 
This reasoning follows the rational choice paradigm [@simonBehavioralModelRational1955], which maintains that behavior is always at least partially influenced by convictions, attitudes, and cost-benefit analyses. 
Also empirically, a study of 1,042 youths from Belgium found that the relation between privacy attitudes and disclosure of personal information was strong (_r_ = .56), whereas the relation between privacy concerns and disclosure was only moderate [_r_ = -.29; @heirmanPredictingAdolescentsDisclosure2013].

@@

For the analyses, we used `r r_citations`.
The code, additional analyses, and a reproducible version of this manuscript can be found on the manuscript's companion website at https://xmtra.github.io/privacy-paradox.

>>>>

However, current research on the privacy paradox has one major limitation. 
To the best of our knowledge, most empirical studies conducted so far have investigated the privacy paradox from a between-person perspective.
By employing empirical tests of relations between people (e.g., cross-sectional questionnaires analyzed with multiple regression or Pearson correlations), studies have analyzed whether people who are more concerned than _others_ also share less personal information than _others_.
Although such a perspective is interesting and represents a viable first step, it cannot make informed claims regarding causality.
The privacy paradox, however, implies a causal perspective:
Does a person, if _he or she_ becomes more concerned about online privacy, then also share less personal information?
This mismatch is problematic because although between-person relations are, except for some special cases, a _necessary_ condition for causal within-person effects, they are by no means a _sufficient_ one.
For example, it could be that the between-person relation is determined by other third variables. 
Hence, as the next step in investigating the privacy paradox and to better understand the causal relation between privacy concerns and information sharing, it is necessary to conduct studies with within-person designs.

@@

Specifically, two of the most influential factors that affect online information sharing are (a) strong subjective norms [@heirmanPredictingAdolescentsDisclosure2013] and (b) expected benefits [@krasnovaOnlineSocialNetworks2010]. 
In other words, users often prioritize social support, special offers, or improved services, accepting that their privacy will be diminished.
Sometimes, privacy concerns do not relate information sharing, because users lack the skills, knowledge, or literacy to change their online behavior, creating feelings of apathy or cynicism [@hargittaiWhatCanReally2016; @hoffmannPrivacyCynicismNew2016].
Likewise, personal information is also often shared by others, a phenomenon described as "networked privacy" [@marwickNetworkedPrivacyHow2014], which further reduces the power of individuals to determine how much personal information can be found online.
@trepteRiskyBehaviorsHow2014 listed several factors that can additionally attenuate the relation:
lack of strength of concerns,
absence of negative personal experiences, 
or situational constraints due to social desirability.
In conclusion, also in the context of the privacy paradox it is _not_ reasonable to expect a _perfect_ relation between attitudes and behaviors.
However, we should still expect to find a relation that is _small_ or _moderate_.

@@

It has been argued that privacy attitudes could bridge the gap between concerns and information sharing [e.g., @dienlinPrivacyParadoxRelic2015]. 
In contrast to general and implicit privacy concerns, privacy attitudes capture a more explicit, specific cognitive appraisal [@tsay-vogelSocialMediaCultivating2018]. 
Because general dispositions oftentimes affect more specific appraisals [@fishbeinPredictingChangingBehavior2010], general concerns about privacy may similarly affect more specific privacy attitudes [@dienlinPrivacyParadoxRelic2015].
This reasoning follows the rational choice paradigm [@simonBehavioralModelRational1955], which maintains that behavior is always at least partially influenced by specific convictions, attitudes, and cost-benefit analyses. 
Therefore, although both variables are related to information disclosure, attitudes are likely the better predictor. 
Also empirically, a study of 1,042 youths from Belgium found that the relation between privacy attitudes and disclosure intentions of personal information was strong (_r_ = .56), whereas the relation between privacy concerns and disclosure intentions was only moderate [_r_ = -.29; @heirmanPredictingAdolescentsDisclosure2013].

@@

For the analyses, we used `r r_citations`.
The code, additional analyses (e.g., ICCs or analyses of invariance), and a reproducible version of this manuscript can be found on the manuscript's companion website at https://xmtra.github.io/privacy-paradox.
